---
title: "Diagnosing breast cancer with k-NN"
date: "`r Sys.time()`"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

## Introduction

Lazy learning is also known as instance-based learning or rote learning. An instance-based learning based learners do not build a model, this method is said to be in a class of non-parametric learning methods-no parameters are learned about the data.

## Data set

The early detection of breast cancer involves examing tissue for abnormal lumps or masses. A clinician examines the cell under amicroscope to determine whether the mass is likely to be the maligant or benign. 
```{r}
wbcd = read.csv("wisc_bc_data.csv", stringsAsFactors = FALSE)
str(wbcd)
```

Delete `id` and see the summary of data:
```{r}
wbcd = wbcd[-1]
table(wbcd$diagnosis)
```

Give `B` and `M` lables:
```{r}
wbcd$diagnosis = factor(wbcd$diagnosis, levels = c("B", "M"), labels = c("Benign", "Maligant"))
round(prop.table(table(wbcd$diagnosis)) * 100, digits = 1)
```

Overview the variables of data
```{r}
summary(wbcd[c("radius_mean", "area_mean", "smoothness_mean")])
```

Normalizing numeric data
```{r}
normalize = function(x){ return((x-min(x))/(max(x)-min(x)))}
wbcd_n = as.data.frame(lapply(wbcd[2:31], normalize))
summary(wbcd_n$area_mean)
```

Create training and testing datasets
```{r}
wbcd_train = wbcd_n[1:469, ]
wbcd_test  = wbcd_n[470:569, ]
wbcd_train_labels = wbcd[1:469, 1]
wbcd_test_labels  = wbcd[470:569, 1]
```

## Training the model
```{r}
if (!require("class")) {install.packages("class")};
library(class)
wbcd_test_pred = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k =21)
```

## Evaluating model performance 
```{r}
if (!require("gmodels")) {install.packages("gmodels")};
library(gmodels)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)
```
From the reult, we can see that only 2 maligant are incorrectly estimated.

## Improving performance
Change standarization
```{r}
wbcd_z = as.data.frame(scale(wbcd[-1:-2]))
summary(wbcd_z$area_mean)
wbcd_train = wbcd_z[1:469, ]
wbcd_test  = wbcd_z[470:569, ]
wbcd_train_labels = wbcd[1:469, 1]
wbcd_test_labels  = wbcd[470:569, 1]
wbcd_test_pred = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)
wbcd_test_pred = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 11)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)
wbcd_test_pred = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 15)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)
```
Alternative value is `k`.



